---
title: "lab_03"
author: "Mitchell Hang"
date: "2023-02-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)
# Load the tidyverse
library(tidyverse)
# Load janitor
library(janitor)
```

## Load data and clean up the column names

Load the two datasets we worked with in the pre_lab: the Maryland payments to vendors and Maryland grants and loans and, using janitor, change the columns so that they are lower-cased and any spaces are replaced with underscores.

```{r}
# Load in the Maryland state agency payments data, making sure that it correctly sets the datatypes using all of the rows
payments <- read_csv("data/State_of_Maryland_Payments_Data__FY2008_to_FY2023.csv", guess_max=322138)
cleaned_payments <- payments %>%
  clean_names()

# Load in the Maryland grants and loan data
md_grants_loans <- read_csv("data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv")
cleaned_md_grants_loans <- md_grants_loans %>%
  clean_names()
```

## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed. Write code to find rows where the vendor name contains the county's school system, starting with a broad search and then narrowing it down. When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find?

A1: I was able to find 60 records.

```{r}
pg_schools <- cleaned_payments %>%
  filter(str_detect(vendor_name, "PRINCE GEORGES COUNTY PUBLIC SCHOOL|PRINCE GEORGES PUBLIC SCHOOL|PRINCE GEORGES COUNTY SCHOOL"))
```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? What do you notice about the years listed?

A2: From what I found, fiscal year 2021 had the largest total amount of $1,470,133,533. I noticed (at least from what I was able to gather) that some of the fiscal years in the dataframe were missing, particularly 2011, 2012, 2014, 2016 and most recently 2022, where their amounts were not documented.

```{r}
pg_schools %>%
  group_by(fiscal_year) %>%
  summarize(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))
```

Q3. You're interested in examining the vendors paid for capital projects at the University of Maryland. Search the payments data to find those payments, then write code to produce a list of each vendor and the total amount it received. Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? Provide at least three examples of vendor names that are similar and should be combined into one. Describe how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?

A3. "WHITING TURNER CONTRACTING CO" looks similar to "WHITING TURNER CONTRACTING COMPANY," "AYERS SAINT GROSS INC" looks similar to "AYERS SAINT GROSS," and "PLANO-COUDON LLC" looks similar to "PLANO-CONDON, LLC." Personally, I would use the "mutate" command to modify some of the vendor names that have the same zip code so that they are consistent (i.e., change "WHITING TURNER CONTRACTING CO" to "WHITING TURNER CONTRACTING COMPANY"), and in the case of GILBANE BUILDING COMPANY, mutate one of the zip codes so that it properly reads "02860" rather than "2860."

```{r}
cleaned_payments %>%
  filter(str_detect(agency_name, "CAPITAL PROJECTS- UNIVERSITY OF MARYLAND")) %>%
  group_by(vendor_name, vendor_zip) %>%
  summarize(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))
```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. Once you find which columns contain the information you want, your code should standardize the data needed to find just the grants for stem cell research in a single filter. Then create a total amount of money for each fiscal year. Write a sentence summarizing how stem cell research funding has changed over time.

A4: 2009 was the fiscal year with the highest total grants for stem cell research funding, with the amount being \$14,643,721. Since then, funding for stem sell research is often around half the 2009 amount for subsequent years, with 2012 being the fiscal year with the second highest amount at \$9,238,365.

```{r}
cleaned_md_grants_loans %>%
  mutate(description = str_to_upper(description)) %>%
  filter(str_detect(description, "STEM CELL RESEARCH GRANT")) %>%
  group_by(fiscal_year) %>%
  summarize(total_amount = sum(amount))
```
