---
title: "pre_lab_01.Rmd"
author: "Mitchell Hang"
date: "2023-02-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

## About this notebook

This notebook contains code and explanatory text that your should review and run as you read through chapters 13 and 15 of the course textbook, "Data Journalism with R and the Tidyverse". Answer questions and edit the document as directed.

Running this notebook will help you understand key data analysis methods and concepts that you will put into practice during this week's lab. When you come across a new command or verb in R, you should add it to your class reference notebook file along with what it's used for.

When you are finished running the code in this notebook, you will push changes to your course GitHub repo, and upload the link to ELMS as instructed.

You also should update your class reference notebook with code examples and what they mean. When you come across a new concept or R verb, put it in your reference notebook along with a sentence about what it does. For example:

```{r}
library(tidyverse)
```

The library function loads an R library such as the tidyverse, providing access to the verbs and functions that library contains. You have to load a library before you can use its functions.

## Instructor in class demonstration outline

-   Load this notebook, and provide overview of R, RStudio (text editor, console, environment window, file explorer), Rmarkdown data notebooks, the reference notebook.
-   Work through the code below, chapter by chapter.
    -   Chapter 13: explain the console and the concept of variables. Explain packages, the Tidyverse, how to install them.
    -   Chapter 15: explain loading libraries, the concept of functions, loading data, the importance of having questions to guide data analysis, and aggregation and summarization.

## Chapter 13

### Task 1: The console

**Task** In the console (look below), type 2+2 and hit enter. What does it return? Type your answer below.

**Answer:** [1] 4

### Task 2: Variables

**Task** Put the following code -- which stores the numeral 2 as a variable called number -- in the console and hit enter.

```{r}
number <- 2
```

Then type this code to multiply the variable called number by itself.

```{r}
number * number
```

What does it return? Type your answer below:

**Answer:** [1] 4

### Task 3: Install packages

**Task** In the console, run the following code to install the Tidyverse:

```{r}
install.packages('tidyverse')
```

**Task** Add that

Then run the following code to install the rmarkdown library

```{r}
install.packages('rmarkdown')
```

### Task 4: Load libraries

**Task** Run the following code in the gray-colored codeblock below -- not in the console -- to load the tidyverse library. To run the code, click the little green play button (left facing arrow) at the top right of the codeblock. In Rmarkdown data notebooks, we write code inside of codeblocks, and explanatory text in the white area outside of it.

```{r}
# This is a comment inside of a codeblock. Comments doesn't run when we run the codeblock.
# Adding this command is a good way to start your class reference notebook.
library(tidyverse)
```

### Task 5: Load data

**Task** Load the UMD courses data by running the following codeblock. The data is contained in an .rds file, which is a binary format for R data.

```{r}
umd_courses <- read_rds("umd_courses.rds")
```

Once it loads, look at the environment window at right in RStudio for umd_courses It's next to a small blue circle with a white arrow in it. To the right of that, look at how many "obs" the dataset has ("obs" is short for observation, which is another way of saying rows or records). Look at how many variables it has (another way of saying columns or fields). Write the number of observations and variables below.

**Answer:** 55,632 observations (rows/records), 7 variables (columns/fields)

### Task 6: Glimpse data

**Task** Run the following codeblock to use glimpse() to get a sense of the data, some of the values in each field/variable/column, and the data types of each field/variable/column.

```{r}
glimpse(umd_courses)
```

In the space below, write the data type for the "id" column and the "seats" column (it's in light gray, offset by \< \>).

**Answer:** id is <chr>, seats is <dbl>

### Task 7: Examine data "head"

**Task** Run the following codeblock to use head() to examine the first six rows of data. Explore the data by clicking the black arrow on the right side of the output table, allowing you to examine additional columns.

```{r}
head(umd_courses)
```

In the space below, write the name of the business from the first row of the "title" column. 

**Answer:** Directed Research

### Task 8: Explore data with interactive dataframe

**Task** look at the environment window at right in RStudio for umd_courses. It's next to a small blue circle with a white arrow in it. Click on the umd_courses -- not the blue circle. It will open the dataframe as an interactive table. Scroll down to see additional rows. Scroll left to see additional columns. Click on a column name to sort by that column.

### Task 9: Group and count

**Task** Run the following codeblock:

```{r}
umd_courses %>%
  group_by(term) %>%
  summarise(
    count_classes = n()
  )
```

Describe what you think each line of this code is doing below, based on the result:

**Answer:** I think these lines of codes are compiling all of the UMD courses taken thus far, sorting them by which term (or semester) they were taken and then counting the number of courses taken during those terms, in chronological order.

### Task 10: Adding arrange to sort the data

**Task** Run the following codeblock:

```{r}
umd_courses %>%
  group_by(term) %>%
  summarise(
    count_classes = n()
  ) %>%
  arrange(desc(count_classes))
```

Compared to the previous codeblock, what has changed and how did that affect the result? 

**Answer:** This time, the terms are sorted by the number of classes students have taken at that time, with the August 2022 term having the most at 4,975 classes.

### Task 11: Grouping by multiple columns

**Task** Run the following codeblock.

```{r}
umd_courses %>%
  group_by(term, department) %>%
  summarise(
    count_classes = n()
  ) %>%
  arrange(term, department)
```

Compared to the previous codeblock, how did the output change and what in the code changed it?

**Answer:** This time, the number of classes taken for each term are split by their departments, starting with the December 2017 term.

### Task 12: Other summary functions

**Task** Run the following codeblock:

```{r}
umd_courses %>%
  group_by(term) %>%
  summarise(
    count_classes = n(),
    total_seats = sum(seats)
  ) %>%
  arrange(desc(total_seats))
```

Compared to the previous codeblock, what has changed and how did that affect the result? 

**Answer:** This time, the terms are sorted both by the number of classes students have taken at that time and the total number of seats, with the August 2022 term having the most at 4,975 classes and 230,796 seats.

### Task 13: Different kinds of summarizations

**Task** Run the following codeblock to group the umd_courses dataframe by term. Then do the following summarizations for each term: count the number of courses, add up the total of all seats, and calculate the mean and median contribution number of seats. Then sort so the term with the most total seats is listed first.

```{r}
umd_courses %>%
  group_by(term) %>%
  summarise(
    count_classes = n(),
    total_seats = sum(seats),
    mean_seats = mean(seats),
    median_seats = median(seats)
  ) %>%
  arrange(desc(total_seats))
```

Compared to the previous codeblock, what has changed and how did that affect the result? What do the top terms have in common?

**Answer:** This time, in addition to the terms being sorted both by the number of classes students have taken at that time and the total number of seats, we also have the mean and median number of seats for those terms. The August 2022 term, which has 4,975 classes and 230,796 seats, has a mean seat number of 46.4 and an average median seat number of 22.

### Task 14: More aggregations

**Task** Run the following codeblock to group the umd_courses dataframe by the department. Then do the following summarizations for each group: count the number of courses, add up the total of all seats, calculate the mean and median number of seats and calculate the smallest and largest seat numbers. Then sort so the department with the largest total number of seats is listed first.

```{r}
umd_courses %>%
  group_by(department) %>%
  summarise(
    count_classes = n(),
    total_seats = sum(seats),
    mean_seats = mean(seats),
    median_seats = median(seats),
    min_seats = min(seats),
    max_seats = max(seats)
  ) %>%
  arrange(desc(total_seats))
```

Compared to the previous codeblock, what has changed and how did that affect the result? How is the data sorted?

**Answer:** This time, we have sorted the classes by department and have them sorted by number of classes taken, total number of seats, as well as the mean, median, minimum and maximum numbers of seats. Business and Management has 1,726 classes and 146,763 total seats, a mean seat number of 85.0, a median seat number of 45, a minimum seat number of 0 and a maximum seat number of 645.

### Task 15: Questions

No code here; I want you to provide three questions that this data suggests to you. What would you like to ask it?

**Answer:**

1.  How many students have taken classes in the Computer Science department during the August 2022 term?

2.  What was the average number of seats for a class offered during the January 2021 term?

3.  What was the maximum number of seats offered by any class from the Linguistics department during any term?
