---
title: "lab_07"
author: "Mitchell Hang"
date: "2023-03-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key if you haven't already. You won't need to install or overwrite it.

```{r echo=FALSE}
census_api_key("3ad46e027fcfceff7f5ae8867c1618c439e44f13")
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2021 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2021 5-year ACS (which is called "acs5") and save them to a dataframe

```{r}
acs <- load_variables(2021, "acs5", cache = TRUE)
View(acs)
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2021 5-year estimates? You will need to find the variable for median age first. 
**A1** The Maryland jurisdiction with the lowest median age is Baltimore City, with a median age of 35.7 years.

```{r}
md_age <- get_acs(geography = "county",
              variables = c(medage = "B01002_001"),
              state = "MD",
              year = 2021)

md_age <- md_age %>% arrange(desc(estimate))

md_age
```

**Q2** Which Maryland counties have a median age of at least 50? You can use the dataframe you just made to answer this, but you must write code to display only those counties with a median age of at least 50. 
**A2** Only two counties have a median age of at least 50; they are Talbot County (51.0 years) and Worcester County (50.4 years).

```{r}
md_age %>%
  filter(estimate >= 50)
```

**Q3** We're interested in knowing more about Montgomery County zip codes where overdose calls are coming from and the demographic characteristics they possess. In particular, we want to calculate a rate of calls per 1,000 population. To do this, you'll need to:

1. Load the Montgomery County 911 overdose calls data.
2. Create a new dataframe with the total number of calls for each zip code.
3. Find the ACS variable pertaining to the total 18 and up population (there are several choices here; you want the simplest one).
4. Get from the ACS data from all Maryland zip codes using that variable.
5. Join the ACS data to the Montgomery County zip code totals dataframe and rename the estimate column to a meaningful name you've given the ACS variable.
6. Add a column to the joined dataframe that calculates a rate of calls per 1,000 population.

Which zip code has the highest rate of calls? Which has the lowest?

**A3** Based on my calculations, the zip code with the highest rate of calls is 20877 in Gaithersburg, with about 1.701 calls for every 1,000 people. The zip code with the lowest rate of calls is 20707 in Laurel, with about 0.040 calls for every 1,000 people. 20889 in Bethesda appears to be the only outlier as it did not give me an estimate of the 18+ population.

```{r}
montgomery_2022_overdoses <- read_csv("data/montgomery_2022_overdoses.csv") %>%
  mutate(zip = as.character(zip))

md_zipcodes <- get_acs(geography="zcta", variables = "B09021_001", state='MD', year=2019)
montgomery_2022_overdoses_acs <- montgomery_2022_overdoses %>% left_join(md_zipcodes, by=c('zip'='GEOID'))
montgomery_2022_overdoses_acs <- montgomery_2022_overdoses_acs %>% mutate(min=(estimate-moe), max=(estimate+moe))

montgomery_2022_overdoses_acs %>%
  group_by(zip, estimate) %>%
  summarize(total_calls = n()) %>%
  mutate(per_capita = (total_calls/estimate)*1000) %>%
  arrange(desc(per_capita))
```

**Q4** Using [CensusReporter.org](https://censusreporter.org/), examine the zip codes with the highest and lowest rates and compare and contrast their demographic profiles. Is there a story here?

**A4** Initially, I didn't know if a story could possibly be conducted because CensusReporter tells me that both zip codes have similar general population amounts (37,094 people for zip 20877 and 36,020 people for zip 20707). If we examine the broader cities, Gaithersburg has a larger general population than Laurel, so it would make sense as to why there would be more calls for every 1,000 people in Gaithersburg compared to Laurel. However, upon examining their zip codes further, we see that the majority of the population in 20707 is Black whereas the majority of the population in 20877 is Hispanic, 20707 has both a larger per capita income and a larger median household income compared to 20877 and there are more people in 20877 that are below the poverty line. Now, I think it might be possible to a story on the rates of overdoses among lower-class residents in Maryland (i.e., those that are below the poverty line) compared to middle and higher-class residents.

**Q5** Go back to your joined dataframe from Q3 and add two more columns representing the minimum and maximum populations based on the estimate and margin of error, then calculate per capita rates for each. Do you get the same two zip codes for the highest and lowest as you did before? Does that give you more or less confidence in your answers to Q3?

**A5** I got the same zip codes as I did for Q3, so I feel more confident in my answers.

```{r}
montgomery_2022_overdoses_acs %>%
  group_by(zip, estimate, min, max) %>%
  summarize(total_calls = n()) %>%
  mutate(per_capita = (total_calls/estimate)*1000, per_capita_min = (total_calls/min)*1000, per_capita_max = (total_calls/max)*1000) %>%
  arrange(desc(per_capita_min))
```

```{r}
montgomery_2022_overdoses_acs %>%
  group_by(zip, estimate, min, max) %>%
  summarize(total_calls = n()) %>%
  mutate(per_capita = (total_calls/estimate)*1000, per_capita_min = (total_calls/min)*1000, per_capita_max = (total_calls/max)*1000) %>%
  arrange(desc(per_capita_max))
```